---
title: "Tidy datasets"
author: "Jay Hesselberth"
date: "December 27, 2015"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---

# Overview

Tidy up all of the R builtin datasets.

```{r library, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)

knitr::opts_chunk$set(comment='#>')

format_strs <- function(strings, sep='.', ...) {
  strings %>%
    stringr::str_to_lower() %>%
    stringr::str_replace_all(' ', sep)
}

```

## ability

```{r ability.cov}
ability.cov_tidy <- ability.cov %>%
  as.data.frame() %>%
  select(-n.obs, -center) %>%
  mutate(test = rownames(.)) %>%
  gather(cov, value, -test) %>%
  tbl_df()

ability.cov_tidy 
```

## airmiles

```{r airmiles}
airmiles_tidy <- airmiles %>%
  as.data.frame() %>%
  setNames('airmiles') %>%
  mutate(year = time(airmiles)) %>%
  tbl_df()

airmiles_tidy
```

## AirPassengers

```{r AirPassengers}
# gawd time series are awful ...
# from http://stackoverflow.com/questions/5331901/transforming-a-ts-in-a-data-frame-and-back

dmn <- list(month.abb, unique(floor(time(AirPassengers))))
AirPassengers_df <- tbl_df(data.frame(matrix(AirPassengers, 12, dimnames = dmn)))

AirPassengers_tidy <- AirPassengers_df %>%
  mutate(month = rownames(.)) %>%
  gather(year, value, -month)

AirPassengers_tidy
```

## airquality

```{r airquality}
# `airquality` already tidy
```

## anscombe

```{r anscombe}
anscombe %>% mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  separate(key, into = c('var','set'), sep = 1) %>%
  spread(var, value) %>%
  arrange(set) -> anscombe_tidy

head(anscombe_tidy)
```

## attenu

```{r attenu}
# `attenu` already tidy
```

## attitude

```{r attitude}
# `attitude` already tidy
```

## austres

```{r austres}
dmn <- list(
  c("Qtr1","Qtr2","Qtr3","Qtr4"),
  unique(floor(time(austres)))
)

austres_tidy <- as.data.frame(t(matrix(c(NA,austres,NA,NA), 4, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(qtr, value, -year) %>%
  arrange(year)

head(austres_tidy)
```

## beavers

```{r beavers}
# Combine the beaver1 and beaver2 datasets ...
beaver1_tidy <- beaver1 %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  mutate(beaver = '1')

beaver2_tidy <- beaver2 %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  mutate(beaver = '2')

beavers_tidy <- rbind_list(beaver1_tidy, beaver2_tidy)

beavers_tidy
```

## BJsales

```{r BJsales}
BJsales_tidy <- data.frame(
  time = as.numeric(time(BJsales)),
  val = as.numeric(BJsales)
)

head(BJsales_tidy)
```

## BJsales.lead

```{r BJsales.lead}
BJsales.lead_tidy <- data.frame(
  time = as.numeric(time(BJsales.lead)),
  val = as.numeric(BJsales.lead)
)

head(BJsales.lead_tidy)
```

## BOD

```{r BOD}
BOD_tidy <- BOD %>% tbl_df %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs)

BOD_tidy
```

## cars

```{r cars}
cars_tidy <- cars %>% tbl_df %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs)

cars_tidy
```

## ChickWeight

```{r ChickWeight, warning=FALSE}
ChickWeight_tidy <- ChickWeight %>%
  gather(key, value, -Chick, convert = TRUE) %>%
  tbl_df

ChickWeight_tidy
```

## chickwts

```{r chickwts}
# `chickwts` already tidy
```

## co2

```{r co2}
dmn <- list(month.abb, unique(floor(time(co2))))

co2_tidy <-  data.frame(t(matrix(co2, 12, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, value, -year) %>%
  arrange(year)

head(co2_tidy)
```

## CO2

```{r CO2}
# `CO2` already tidy
```

## crimtab

```{r crimtab}
crimtab_tidy <- as.data.frame(crimtab)

head(crimtab_tidy)
```

## discoveries

```{r discoveries}
discoveries_tidy <- data.frame(
  year = as.numeric(time(discoveries)),
  count = as.numeric(discoveries))

head(discoveries_tidy)
```

## DNase

```{r DNase}
# `DNase` already tidy
```

## esoph

```{r esoph}
# `esoph` already tidy
```

## euro

```{r euro}
euro_tidy <- data.frame(
  Country = names(euro),
  val = as.numeric(euro))

head(euro_tidy)
```

## euro.cross

```{r euro.cross}
euro.cross_tidy <- as.data.frame(euro.cross) %>%
  mutate(Euro_curr = rownames(euro.cross)) %>%
  gather(Euro_curr2, value, -Euro_curr) %>%
  arrange(Euro_curr)

head(euro.cross_tidy)
```

## eurodist

```{r eurodist}
# `eurodist` is a dist object (and I don't feel like figuring it out)
```

## EuStockMarkets

```{r EuStockMarkets}
dmn <- list(colnames(EuStockMarkets), time(EuStockMarkets))

EuStockMarkets_tidy <- data.frame(t(matrix(EuStockMarkets, 4, dimnames = dmn))) %>%
  mutate(time_ind = rownames(.)) %>%
  gather(key, val, -time_ind) %>%
  separate(time_ind, into = c("year","time"), sep = 4) %>%
  arrange(year)

head(EuStockMarkets_tidy)
```

## faithful

```{r faithful}
# `faithful` already tidy
```

## fdeaths

```{r fdeaths}
dmn <- list(month.abb, unique(floor(time(fdeaths))))

fdeaths_tidy <- data.frame(t(matrix(fdeaths, 12, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, deaths, -year) %>%
  arrange(year)

head(fdeaths_tidy)
```

## Formaldehyde

```{r Formaldehyde}
# `Formaldehyde` already tidy
```

## freeny freeny.x freeny.y

```{r freeny freeny.x freeny.y}
freeny_y_df <- as.data.frame(t(matrix(c(NA,freeny.y), 4)))

colnames(freeny_y_df) <- c("Qtr1","Qtr2","Qtr3","Qtr4")
rownames(freeny_y_df) <- 1962:1971

freeny_y_tmp <- freeny_y_df %>% mutate(year = rownames(freeny_y_df)) %>%
  gather(Qtr, value, -year) %>%
  arrange(year)

freeny_y_tidy <- freeny_y_tmp[2:40,]

freeny_tidy <- freeny %>% mutate(year = freeny_y_tidy$year, Qtr = freeny_y_tidy$Qtr) %>%
  gather(exp_var, value, -year, -Qtr)

head(freeny_tidy)
```

## HairEyeColor

```{r HairEyeColor}
males <- as.data.frame(HairEyeColor[1:4,1:4,1])

males_tidy <- males %>% 
  mutate(hair_color = rownames(males), sex = rep("male",4)) %>%
  gather(eye_color, value, -sex, -hair_color)

females <- as.data.frame(HairEyeColor[1:4,1:4,2])

females_tidy <- females %>% 
  mutate(hair_color = rownames(females), sex = rep("female",4)) %>%
  gather(eye_color, value, -sex, -hair_color)

HairEyeColor_tidy <- rbind(males_tidy,females_tidy)
HairEyeColor_tidy <- HairEyeColor_tidy[,c("sex","eye_color","hair_color","value")]

head(HairEyeColor_tidy)
```

## Harman23.cor

## Harman 74.cor

## Indometh

```{r Indometh}
# `Indometh` already tidy
```

## infert

```{r infert}
# `infert` already tidy
```

## InsectSprays

```{r InsectSprays}
# `InsectSprays` already tidy
```

## iris

```{r iris}
# `iris` already tidy
```

## iris3

```{r iris3}
iris3_tidy <- as.data.frame(rbind(iris3[,,1],iris3[,,2],iris3[,,3])) %>%
  mutate(species = c(
    rep(dimnames(iris3)[[3]][1], length(iris3[,1,1])),
    rep(dimnames(iris3)[[3]][2], length(iris3[,1,2])),
    rep(dimnames(iris3)[[3]][3], length(iris3[,1,3]))
  ))

head(iris3_tidy)
```

## islands

```{r islands}
islands_tidy <- data.frame(
  land = names(islands),
  area = as.numeric(islands)
)

head(islands_tidy)
```

## JohnsonJohnson

```{r JohnsonJohnson}
dmn <- list(
  c("Qtr1","Qtr2","Qtr3","Qtr4"),
  unique(floor(time(JohnsonJohnson)))
)

JohnsonJohnson_tidy <- as.data.frame(t(matrix(JohnsonJohnson, 4, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(Qtr, val, -year) %>%
  arrange(year)

head(JohnsonJohnson_tidy)
```

## LakeHuron

```{r LakeHuron}
LakeHuron_tidy <- data.frame(
  year = unique(time(LakeHuron)),
  level = as.numeric((LakeHuron))
)

head(LakeHuron_tidy)
```

## ldeaths (mdeaths, fdeaths)

```{r ldeaths mdeaths fdeaths}
dmn <- list(month.abb, unique(floor(time(ldeaths))))

ldeaths_tidy <- data.frame(t(matrix(ldeaths, 12, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, deaths, -year) %>%
  arrange(year)

head(ldeaths_tidy)
```

## lh

```{r lh}
lh_tidy <- data.frame(
  time_int = as.numeric(time(lh)),
  lh = as.numeric(lh)
)

head(lh_tidy)
```

## LifeCycleSavings

```{r LifeCycleSavings}
LifeCycleSavings_tidy <- LifeCycleSavings %>%
  mutate(country = rownames((LifeCycleSavings))) %>%
  gather(stat, val, -country) %>%
  arrange(country)

head(LifeCycleSavings_tidy)
```

## Loblolly

```{r Loblolly}
Loblolly_tidy <- as.data.frame(Loblolly) %>%
  mutate(index = as.numeric(rownames(Loblolly))) %>%
  arrange(index)

head(Loblolly_tidy)
```

## longley

```{r longley}
# `longley` already tidy
```

## lynx

```{r lynx}
lynx_tidy <- data.frame(
  year = as.numeric(time(lynx)),
  val = as.numeric(lynx)
)

head(lynx_tidy)
```

## mdeaths
```{r mdeaths}
dmn <- list(month.abb, unique(floor(time(mdeaths))))

mdeaths_tidy <- data.frame(t(matrix(mdeaths, 12, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, deaths, -year) %>%
  arrange(year)

head(mdeaths_tidy)
```

## morley

```{r morley}
# `morley` already tidy
```

## mtcars

```{r mtcars}
mtcars_tidy <- mtcars %>%
  mutate(car = format_strs(rownames(.))) %>% 
  gather(key, value, -car) %>% tbl_df

mtcars_tidy
```

## nhtemp

```{r nhtemp}
nhtemp_tidy <- data.frame(
  year = as.numeric(time(nhtemp)),
  val = as.numeric(nhtemp)
)

head(nhtemp_tidy)
```

## Nile

```{r Nile}
Nile_tidy <- data.frame(
  year = as.numeric(time(Nile)),
  val = as.numeric(Nile)
)

head(Nile_tidy)
```

## nottem

```{r nottem}
dmn <- list(month.abb, unique(floor(time(nottem))))

nottem_tidy <- data.frame(t(matrix(nottem,12,dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, val, -year) %>%
  arrange(year)

head(nottem_tidy)
```

## npk
```{r npk}
# `npk` already tidy
```

## occupationalstatus

```{r occupationalstatus}
occupationalStatus_tidy <- as.data.frame(occupationalStatus) %>%
  arrange(origin)

head(occupationalStatus_tidy)
```

## Orange

```{r Orange}
# `Orange` already tidy
```

## OrchardSprays

```{r OrchardSprays}
# `OrchardSprays` already tidy
```

## PlantGrowth

```{r PlantGrowth}
# `PlantGrowth` already tidy
```

## precip

```{r precip}
precip_tidy <- data.frame(
  City = names(precip),
  val = as.numeric(precip)
) %>% arrange(City)

head(precip_tidy)
```

## presidents

```{r presidents}
dmn <- list(
  c("Qtr1","Qtr2","Qtr3","Qtr4"),
  unique(floor(time(presidents)))
)

presidents_tidy <- data.frame(t(matrix(presidents,4,dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(Qtr, val, -year) %>%
  arrange(year)

head(presidents_tidy)
```

## pressure

```{r pressure}
# `pressure` already tidy
```

## Puromycin

```{r Puromycin}
# `Puromycin` already tidy
```

## quakes

```{r quakes}
quakes_tidy <- quakes %>% mutate(event = rownames(.))

head(quakes_tidy)
```

## randu

```{r randu}
# `randu` already tidy
```

## rivers

```{r rivers}
rivers_tidy <- data.frame(index = 1:length(rivers), length = rivers)

head(rivers_tidy)
```

## rock

```{r rock}
# `rock` already tidy
```

## Seatbelts

```{r Seatbelts}
Seatbelts_tidy <- data.frame(Seatbelts) %>%
  mutate(year = as.numeric(floor(time(Seatbelts))),
         month = rep(month.abb,(length(as.numeric(floor(time(Seatbelts))))/12))) %>%
  gather(key, val, -year,-month) %>%
  arrange(year)

head(Seatbelts_tidy)
```

## sleep

```{r sleep}
# `sleep` already tidy
```

## stackloss

```{r stackloss}
# `stackloss` already tidy
```

## state

```{r state.abb state.name state.area state.center state.division}
state_tidy <- data.frame(state = state.name, state_abb = state.abb, area = state.area,
                         center_x = state.center$x, center_y = state.center$y, division = state.division)

head(state_tidy)
```

## sunspot.month

```{r sunspot.month}
dmn <- list(month.abb, unique(floor(time(sunspot.month))))

sunspot.month_tidy <- data.frame(t(matrix(c(sunspot.month,NA,NA,NA),12,dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, val, -year) %>%
  arrange(year)

head(sunspot.month_tidy)
```

## sunspot.year

```{r sunspot.year}
sunspot.year_tidy <- data.frame(
  year = time(sunspot.year),
  val = as.numeric(sunspot.year)
)

head(sunspot.year_tidy)
```

## sunspots

```{r sunspots}
dmn <- list(month.abb, unique(floor(time(sunspots))))

sunspots_tidy <- data.frame(t(matrix(sunspots,12,dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month, val, -year) %>%
  arrange(year)

head(sunspots_tidy)
```

## swiss

```{r swiss}
swiss_tidy <- swiss %>% mutate(province = rownames(.)) %>%
  gather(key, val, -province) %>%
  arrange(province)

head(swiss_tidy)
```

## Theoph

```{r Theoph}
# `Theoph` already tidy
```

## Titanic

```{r Titanic}
Titanic_tidy <- data.frame(NULL)

for(i in 1:2){
  for(j in 1:2){
    x <- as.data.frame(Titanic[,,i,j]) %>%
      mutate(class = rownames(.),
             age = rep(dimnames(Titanic)$Age[i],length(Titanic[,1,i,j])),
             survived = rep(dimnames(Titanic)$Survived[j],length(Titanic[,1,i,j])))
    Titanic_tidy <- rbind(Titanic_tidy,x)
  }
}

head(Titanic_tidy)
```

## ToothGrowth

```{r ToothGrowth}
ToothGrowth_tidy <- ToothGrowth %>%
  mutate(index = as.numeric(rownames(.))) %>%
  gather(key,val,-index) %>%
  arrange(index)

head(ToothGrowth_tidy)
```

## treering

```{r treering}
treering_tidy <- data.frame(
  time = time(treering),
  val = as.numeric(treering)
)

head(treering_tidy)
```

## trees
```{r trees}
trees_tidy <- trees %>% mutate(index = as.numeric(rownames(.))) %>%
  gather(key, val, -index) %>%
  arrange(index)

head(trees_tidy)
```

## UCBAdmissions

```{r UCBAdmissions}
UCBAdmissions_tidy <- data.frame(NULL)

for(i in 1:length(UCBAdmissions[1,1,])){
  x <- as.data.frame(UCBAdmissions[,,i]) %>%
    mutate(admit = rownames(.),
           dept = rep(dimnames(UCBAdmissions)$Dept[i],length(UCBAdmissions[,1,i]))) %>%
    gather(sex,val,-admit,-dept)
  
  UCBAdmissions_tidy <- rbind(UCBAdmissions_tidy,x)
}

head(UCBAdmissions_tidy)
```

## UKDriverDeaths

```{r UKDriverDeaths}
dmn <- list(month.abb, unique(floor(time(UKDriverDeaths))))

UKDriverDeaths_tidy <- data.frame(t(matrix(UKDriverDeaths,12,dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(month,val,-year) %>%
  arrange(year)

head(UKDriverDeaths_tidy)
```

## UKgas

```{r UKgas}
dmn <- list(
  c("Qtr1","Qtr2","Qtr3","Qtr4"),
  unique(floor(time(UKgas)))
)

UKgas_tidy <- data.frame(t(matrix(UKgas, 4, dimnames = dmn))) %>%
  mutate(year = rownames(.)) %>%
  gather(Qtr,val,-year) %>%
  arrange(year)

head(UKgas_tidy)
```

## USArrests

```{r USArrests}
USArrests_tidy <- as.data.frame(USArrests) %>%
  mutate(state = rownames(.)) %>%
  gather(type, val, -state) %>%
  arrange(state)

head(USArrests_tidy)
```

## UScitiesD
```{r UScitiesD}
# `UScitiesD` is a dist file
```

## USJudgeRatings

```{r USJudgeRatings}
USJudgeRatings_tidy <- as.data.frame(USJudgeRatings) %>%
  mutate(judge = rownames(.)) %>%
  gather(var, val, -judge) %>%
  arrange(judge)

head(USJudgeRatings_tidy)
```

## USPersonalExpenditures

```{r USPersonalExpenditures}
USPersonalExpenditure_tidy <- as.data.frame(USPersonalExpenditure) %>%
  mutate(type = rownames(.)) %>%
  gather(year, val, -type) %>%
  arrange(year)

head(USPersonalExpenditure_tidy)
```

## uspop

```{r uspop}
uspop_tidy <- data.frame(
  time = time(uspop),
  val = as.numeric(uspop)
)

head(uspop_tidy)
```

## VADeaths

```{r VADeaths}
VADeaths_tidy <- as.data.frame(VADeaths) %>%
  mutate(age = rownames(.)) %>%
  gather(key, val, -age) %>%
  arrange(age)

head(VADeaths_tidy)
```

## volcano

```{r volcano}
# `volcano` is a topographic map
```

## warpbreaks

```{r warpbreaks}
# `warpbreaks` already tidy
```

## women

```{r women}
# `women` already tidy
```

## World Phones

```{r WorldPhones}
WorldPhones_tidy <- WorldPhones %>%
  as.data.frame() %>%
  mutate(year = rownames(.)) %>%
  gather(continent, value, -year) %>% tbl_df

WorldPhones_tidy
```

## WWWusage

```{r WWWusage}
WWWusage_tidy <- data.frame(
  time = time(WWWusage),
  val = as.numeric(WWWusage)
)

head(WWWusage_tidy)
```