---
title: "Tidy datasets"
author: "Jay Hesselberth"
date: "December 27, 2015"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---

# Overview

Tidy up all of the R builtin datasets.

```{r library, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(tidytime)

knitr::opts_chunk$set(comment='#>')

format_strs <- function(strings, sep='.', ...) {
  strings %>%
    stringr::str_to_lower() %>%
    stringr::str_replace_all(' ', sep)
}

```

## ability

```{r ability.cov}
ability.cov_tidy <- ability.cov %>%
  as.data.frame() %>%
  select(-n.obs, -center) %>%
  mutate(test = rownames(.)) %>%
  gather(cov, value, -test) %>%
  separate(cov, c("cov", "cat")) %>%
  select(-cov) %>%
  tbl_df()

ability.cov_tidy 
```

## airmiles

```{r airmiles}
airmiles_tidy <- airmiles %>%
  as.data.frame() %>%
  setNames('airmiles') %>%
  mutate(year = time(airmiles)) %>%
  tbl_df()

airmiles_tidy
```

## AirPassengers

```{r AirPassengers}
# gawd time series are awful ...
# from http://stackoverflow.com/questions/5331901/transforming-a-ts-in-a-data-frame-and-back

dmn <- list(month.abb, unique(floor(time(AirPassengers))))
AirPassengers_df <- tbl_df(data.frame(matrix(AirPassengers, 12, dimnames = dmn)))

AirPassengers_tidy <- AirPassengers_df %>%
  mutate(month = rownames(.)) %>%
  gather(year, value, -month) %>%
  separate(year, c("x", "year"), sep = 1) %>%
  select(-x)

AirPassengers_tidy
```

## airquality

```{r airquality}
# 'airquality' is already tidy
head(airquality)
```

## anscombe

```{r anscombe}
anscombe_tidy <- anscombe %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  separate(key, into = c("var", "set"), sep=1) %>%
  spread(var, value) %>%
  arrange(set, obs)

head(anscombe_tidy)
```

## austres

```{r austres}
austres_tidy <- austres %>%
  tidytime() %>%
  setNames(c("year", "series","residents")) %>%
  select(-series)
austres_tidy
```

## attenu

```{r attenu}
# `attenu` already tidy
head(attenu)
```

## attitude

```{r attitude}
# 'attitude' is already tidy
head(attitude)
```

## beavers

```{r beavers}
# Combine the beaver1 and beaver2 datasets ...
beaver1_tidy <- beaver1 %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  mutate(beaver = '1')

beaver2_tidy <- beaver2 %>%
  mutate(obs = seq_len(n())) %>%
  gather(key, value, -obs) %>%
  mutate(beaver = '2')

beavers_tidy <- rbind_list(beaver1_tidy, beaver2_tidy)

beavers_tidy
```

## BJsales

```{r BJsales}
BJsales_tidy <- BJsales %>%
  tidytime() %>%
  select(-series)
BJsales_tidy
```

## BJsales.lead

```{r BJsales.lead}
BJsales.lead_tidy <- BJsales.lead%>%
  tidytime() %>%
  select(-series)
BJsales.lead_tidy
```

## BOD

```{r BOD}
# 'BOD' is already tidy
head(BOD)
```

## cars

```{r cars}
# 'cars' is already tidy
head(cars)
```

## ChickWeight

```{r ChickWeight, warning=FALSE}
ChickWeight_tidy <- ChickWeight %>%
  gather(key, value, -Chick, convert = TRUE) %>%
  tbl_df

ChickWeight_tidy
```

## chickwts

```{r chickwts}
# `chickwts` already tidy
head(chickwts)
```

## CO2

```{r CO2}
# `CO2` already tidy
head(CO2)
```

## co2

```{r co2}
co2_tidy <- co2 %>%
  tidytime() %>%
  setNames(c("year", "series", "co2")) %>%
  select(-series)
co2_tidy
```

## crimtab

```{r crimtab}
crimtab_tidy <- crimtab %>%
  as.data.frame() %>%
  setNames(c("finger_len", "bod_height","count")) %>%
  tbl_df()
crimtab_tidy
```

## discoveries

```{r discoveries}
discoveries_tidy <- discoveries %>%
  tidytime() %>%
  setNames(c("year", "series", "disc")) %>%
  select(-series)
discoveries_tidy
```

## DNase

```{r DNase}
# `DNase` already tidy
head(DNase)
```

## esoph

```{r esoph}
# `esoph` already tidy
head(esoph)
```

## euro

```{r euro}
euro_tidy<- euro %>%
  data_frame(country = names(euro)) %>%
  setNames(c("rate", "country"))
head(euro_tidy)
```

## euro.cross

```{r euro.cross}
euro.cross_tidy <- euro.cross %>%
  as.data.frame() %>%
  mutate(from = row.names(.)) %>%
  gather(to, value, -from) %>%
  arrange(from, to) 
head(euro.cross_tidy)
```

## eurodist

```{r eurodist}
eurodist_tidy <- eurodist %>%
  as.matrix() %>%
  as.data.frame() %>%
  mutate("from" = row.names(.)) %>%
  gather(to, dist, -from) %>%
  arrange(from)
head(eurodist_tidy)
```

## EuStockMarkets

```{r EuStockMarkets}
EuStockMarkets_tidy <- EuStockMarkets %>%
  tidytime() %>%
  setNames(c("year", "country", "value"))
EuStockMarkets_tidy
```

## faithful

```{r faithful}
# `faithful` already tidy
head(faithful)
```

## fdeaths

```{r fdeaths}
fdeaths_tidy <- fdeaths %>%
  tidytime() %>%
  select(-series) %>%
  setNames(c("year", "deaths"))
fdeaths_tidy  
```

## Formaldehyde

```{r Formaldehyde}
# `Formaldehyde` already tidy
head(Formaldehyde)
```

## freeny

```{r freeny}
# freeny is tidy version of freeny.x+freeny.y
head(freeny)
```

## freeny.x

```{r freeny.x}
#see freeny

```

## freeny.y

```{r freeny.y}
# see freeny

```

## HairEyeColor

```{r HairEyeColor}
HairEyeColor_tidy <- HairEyeColor %>%
  as.data.frame() %>%
  tbl_df()
HairEyeColor_tidy
```

## Harman23.cor

```{r Harman23.cor}
Harman23_tidy <- Harman23.cor %>%
  as.data.frame() %>%
  mutate(measurement = row.names(.)) %>%
  gather(compare, value, -measurement) %>%
  arrange(measurement, compare)
head(Harman23_tidy)
```

## Harman74.cor

```{r Harman74.cor}
Harman74_tidy <- Harman74.cor %>%
  as.data.frame() %>%
  mutate(measurement = row.names(.)) %>%
  gather(compare, value, -measurement) %>%
  arrange(measurement, compare)
head(Harman74_tidy)
```

## Indometh
```{r Indometh}
# 'Indometh' already tidy
head(Indometh)
```

### infert
```{r infert}
# 'infert' already tidy
head(infert)
```

## InsectSprays
```{r InsectSprays}
# `InsectSprays` already tidy
head(InsectSprays)
```

## iris
```{r iris}
# 'iris' is already tidy
head(iris)
```

## iris3

```{r iris3}
iris3_tidy <- iris3 %>%
  as.data.frame()
head(iris3_tidy)

```

## islands

```{r islands}
islands_tidy <- islands %>%
  as.data.frame() %>%
  setNames("area") %>%
  mutate(island = row.names(.)) %>%
  select(2,1) %>%
  tbl_df()
islands_tidy
```

## JohnsonJohnson

```{r JohnsonJohnson}
JohnsonJohnson_tidy <- JohnsonJohnson %>%
  tidytime() %>%
  select(-series) %>%
  setNames(c("year", "sales"))
JohnsonJohnson_tidy
```

## LakeHuron

```{r LakeHuron}
LakeHuron_tidy <- LakeHuron %>%
  tidytime() %>%
  setNames(c("year", "series", "level_ft")) %>%
  select(-series)
LakeHuron_tidy
```

## ldeaths

```{r ldeaths}
ldeaths_tidy <- ldeaths %>%
  tidytime() %>%
  setNames(c("year", "series", "deaths"))
ldeaths_tidy
```

## lh

```{r lh}
lh_tidy <- lh %>%
  tidytime() %>%
  setNames(c("time_min", "series", "lh")) %>%
  select(-series)
lh_tidy
```

## LifeCycleSavings
```{r LifeCycleSavings}
LifeCycleSavings_tidy <- LifeCycleSavings %>%
  mutate(country = row.names(.)) %>%
  select(6,1,2,3,4,5) %>%
  tbl_df()
LifeCycleSavings_tidy
```

## Loblolly
```{r Loblolly}
Loblolly_tidy <- Loblolly %>%
  mutate(tree_num = row.names(.)) %>%
  select(4,1,2,3) %>%
  tbl_df()
Loblolly_tidy
```

## longley
```{r longley}
# longley is already tidy
head(longley)
```

## lynx

```{r lynx}
lynx_tidy <- lynx %>%
  tidytime() %>%
  setNames(c("year", "series", "trappings")) %>%
  select(-series)
lynx_tidy
```

## mdeaths

```{r mdeaths}
mdeaths_tidy <- mdeaths %>%
  tidytime() %>%
  setNames(c("year", "series", "deaths")) %>%
  select(-series)
mdeaths_tidy
```

## morley
```{r morley}
# morley is already tidy
head(morley)
```

## mtcars
```{r mtcars}
mtcars_tidy <- mtcars %>%
  mutate(car = row.names(.)) %>%
  select(12,1:11) %>%
  tbl_df()
mtcars_tidy
```

## nhtemp

```{r nhtemp}
nhtemp_tidy <- nhtemp %>%
  tidytime() %>%
  setNames(c("year", "series", "temp_F")) %>%
  select(-series)
nhtemp_tidy
```

## Nile

```{r Nile}
Nile_tidy <- Nile %>%
  tidytime() %>%
  setNames(c("year", "series", "flow")) %>%
  select(-series)
Nile_tidy
```

## nottem

```{r nottem}
nottem_tidy <- nottem %>%
  tidytime() %>%
  setNames(c("year", "series", "temp_F")) %>%
  select(-series)
nottem_tidy
```

## npk
```{r npk}
# npk is already tidy
head(npk)
```

## occupationalStatus

```{r occupationalStatus}
occupationalStatus_tidy <- occupationalStatus %>%
  as.data.frame() %>%
  setNames(c("father_occ", "son_occ", "count")) %>%
  arrange(father_occ)
head(occupationalStatus_tidy)
```

## Orange
```{r Orange}
# 'Orange' already tidy
head(Orange)
```

## OrchardSprays
```{r OrchardSrays}
# 'OrchardSprays' alreayd tidy
head(OrchardSprays)
```

## PlantGrowth
```{r PlantGrowth}
# 'PlantGrowth' already tidy
head(PlantGrowth)
```

## precip

```{r precip}
precip_tidy <- precip %>%
  data.frame("city" = names(precip), "precip" = as.numeric(precip)) %>%
  select(city, precip) %>%
  arrange(city)
head(precip_tidy)
```

## presidents

```{r presidents}
presidents_tidy <- presidents %>%
  tidytime() %>%
  setNames(c("year", "series", "rating")) %>%
  select(-series)
presidents_tidy
```

## pressure
```{r pressure}
# 'pressure' is already tidy
head(pressure)
```

## Puromycin
```{r Puromycin}
# 'Puromycin' is already tidy
head(Puromycin)
```

## quakes
```{r quakes}
# 'quakes' is already tidy
head(quakes)
```

## randu
```{r randu}
# 'randu' is already tidy
head(randu)
```

## rivers

```{r rivers}
rivers_tidy <- rivers %>%
  as.data.frame() %>%
  setNames("length_mi") %>%
  mutate(river_num = seq_len(n())) %>%
  select(2,1) %>%
  tbl_df()
rivers_tidy
```

## rock
```{r rock}
rock_tidy <- rock %>%
  mutate(samp_num = seq_len(n())) %>%
  select(5,1:4) %>%
  tbl_df()
rock_tidy
```

## sleep
```{r sleep}
# 'sleep' is already tidy
head(sleep)
```

## stack.loss

```{r stack.loss}
#see stackloss
```

## stack.x

```{r stack.x}
# see stackloss
```

## stackloss
```{r stackloss}
# 'stackloss' already tidy (stack.loss + stack.x)
head(stackloss)
```

## state.abb

```{r state.abb}
# see states
```

## state.area

```{r state.area}
# see states
```

## state.center

```{r state.center}
# see states
```

## state.division

```{r state.division}
# see states
```

## state.name

```{r state.name}
# see states
```

## state.region

```{r state.region}
# see states
```

## state.x77

```{r state.x77}
# see states
```

## states

```{r states}
state.abb_tidy <- as.data.frame(state.abb) 
state.area_tidy <- as.data.frame(state.area)
state.center_tidy <- as.data.frame(state.center)
state.division_tidy <- as.data.frame(state.division)
state.name_tidy <- as.data.frame(state.name)
state.region_tidy <- as.data.frame(state.region)
state.x77_tidy <- state.x77 %>%
  as.data.frame() %>%
  mutate(state = row.names(.)) %>%
  select(9,1:8)

states_tidy <- cbind(state.abb_tidy, state.area_tidy, state.center_tidy, state.division_tidy, state.region_tidy, state.x77_tidy) %>%
  tbl_df()
states_tidy
```

## sunspot.month

```{r sunspot.month}
sunspot.month_tidy <- sunspot.month %>%
  tidytime() %>%
  setNames(c("year", "series", "sunspots")) %>%
  select(-series)
sunspot.month_tidy
```

## sunspot.year

```{r sunspot.year}
sunspot.year_tidy <- sunspot.year %>%
  tidytime() %>%
  setNames(c("year", "series", "sunspots")) %>%
  select(-series)
sunspot.year_tidy
```

## sunspots

```{r sunspots}
sunspots_tidy <- sunspots %>%
  tidytime() %>%
  setNames(c("year", "series", "sunspots")) %>%
  select(-series)
sunspots_tidy
```

## swiss
```{r swiss}
swiss_tidy <- swiss %>%
  mutate(region = rownames(.)) %>%
  select(7,1:6) %>%
  tbl_df()
swiss_tidy
```

## Theoph
```{r Theoph}
# 'Theoph' is already tidy
head(Theoph)
```

## Titanic

```{r Titanic}
Titanic_tidy <- Titanic %>%
  as.data.frame() %>%
  tbl_df()
Titanic_tidy
```

## ToothGrowth
```{r ToothGrowth}
# 'ToothGrowth' is already tidy
head(ToothGrowth)
```

## treering

```{r treering}
treering_tidy <- treering %>%
  tidytime() %>%
  setNames(c("year", "series", "width")) %>%
  select(-series)
treering_tidy
```

## trees
```{r trees}
trees_tidy <- trees %>%
  mutate(tree_num = seq_len(n())) %>%
  select(4,1:3) %>%
  tbl_df()
trees_tidy
```

## UCBAdmissions

```{r USBAdmissions}
UCBAdmissions_tidy <- UCBAdmissions %>%
  as.data.frame() %>%
  tbl_df()
UCBAdmissions_tidy
```

## UKDriverDeaths

```{r UKDriverDeaths}
UKDriverDeaths_tidy <- UKDriverDeaths %>%
  tidytime() %>%
  setNames(c("year", "series", "deaths")) %>%
  select(-series)
UKDriverDeaths_tidy
```

## UKgas

```{r UKgas}
UKgas_tidy <- UKgas %>%
  tidytime() %>%
  setNames(c("year", "series", "gas")) %>%
  select(-series)
UKgas_tidy
```

## USAccDeaths

```{r USAccDeaths}
USAccDeaths_tidy <- USAccDeaths %>%
  tidytime() %>%
  setNames(c("year", "series", "deaths")) %>%
  select(-series)
USAccDeaths_tidy
```

## USArrests
```{r USArrests}
USArrests_tidy <- USArrests %>%
  mutate(state = rownames(.)) %>%
  select(5,1:4) %>%
  tbl_df()
USArrests_tidy
```

## UScitiesD

```{r UScitiesD}
UScitiesD_tidy <- UScitiesD %>%
  as.matrix() %>%
  as.data.frame() %>%
  mutate("from" = row.names(.)) %>%
  gather(to, dist, -from) %>%
  arrange(from)
head(UScitiesD_tidy)
```

## USJudgeRatings
```{r USJudgeRatings}
USJudgeRatings_tidy <- USJudgeRatings %>%
  mutate(judge= rownames(.)) %>%
  select(13,1:12) %>%
  tbl_df()
USJudgeRatings_tidy
```

## USPersonalExpenditure
```{r USPersonalExpenditure}
USPersonalExpenditure_tidy <- USPersonalExpenditure %>%
  as.data.frame() %>%
  mutate(group = rownames(.)) %>%
  gather(year, expense, -group) %>%
  tbl_df()
USPersonalExpenditure_tidy
```

## uspop

```{r uspop}
uspop_tidy <- uspop %>%
  tidytime() %>%
  setNames(c("year", "series", "pop")) %>%
  select(-series)
uspop_tidy
```

## VADeaths
```{r VADeaths}
VADeaths_tidy <- VADeaths %>%
  as.data.frame() %>%
  mutate(age = rownames(.)) %>%
  gather(demographic, deaths, -age) %>%
  tbl_df()
VADeaths_tidy
```

## volcano
```{r volcano}
volcano_tidy <- volcano %>%
  as.data.frame() %>%
  mutate(x = row.names(.)) %>%
  gather(y, topo, -x) %>%
  separate("y", c("junk", "y"), sep = 1) %>%
  select(-junk)
head(volcano_tidy)
```

## warpbreaks
```{r warpbreaks}
# 'warpbreaks' is already tidy
head(warpbreaks)
```

## women
```{r women}
# 'women' is already tidy
head(women)
```

## WorldPhones

```{r WorldPhones}
WPs_tidy <- WorldPhones %>% 
  as.data.frame() %>%
  mutate(year = rownames(.)) %>%
  gather(continent, value, -year) %>%
  tbl_df()
WPs_tidy
```

## WWWusage

```{r WWWusage}
WWWusage_tidy <- WWWusage %>%
  tidytime() %>%
  setNames(c("time", "series", "users")) %>%
  select(-series)
WWWusage_tidy
```
